---
title: "Project week 2"
author: "Quan Luong"
date: "May 24, 2017"
output: html_document
---

#Q1: Loading and preprocessing the data
```{r}
setwd("D:/TIKI/Quan Luong/Coursera/Course 5_Reproducible Research")
# load raw data
data <- read.csv("./repdata_data_activity/activity.csv")
# re-format date 
data$date <- as.POSIXct(data$date, format="%Y-%m-%d")
```

#Q2: What is mean total number of steps taken per day?

```{r, warning=FALSE, message=FALSE}
library(dplyr)
data_na_rm <- na.omit(data)
data_sum <- data_na_rm %>% 
            group_by(date) %>% 
            summarise(sum_step = sum(steps))
hist(data_sum$sum_step, xlab="Total number of steps", 
     main="Histogram of the total number of steps taken each day")
```

```{r, echo=FALSE}
print(paste0("mean total number of steps taken per day ", round(mean(data_sum$sum_step),2)))
print(paste0("median total number of steps taken per day ", round(median(data_sum$sum_step),2)))

```

#Q3: What is the average daily activity pattern?

```{r}
library(ggplot2)
data_interval <- data_na_rm %>% 
                 group_by(interval) %>% 
                 summarise(sum_step = sum(steps),
                           mean_step = mean(steps))
ggplot(data_interval, aes(interval, sum_step)) + geom_line() + 
     xlab("interval (minutes)") + 
     ylab("Average number of steps") + 
     ggtitle("Time-series of the average number of steps per intervals\n(NA removed)")
```

```{r, echo=FALSE}
print(paste0("Averaged number of steps peaks at interval ", data_interval$interval[data_interval$mean_step == max(data_interval$mean_step)]))
```

#Q4: Imputing missing values

## Report the total number of missing values in the dataset
```{r, echo=FALSE}
print(paste0("Total number of missing values in the dataset ",sum(is.na(data))))
```

## Filling missing data
```{r}
data_na_filled <- data
data_na_filled$steps <- ifelse(!is.na(data_na_filled$steps), 
                               data_na_filled$steps, 
                               ifelse(data_na_filled$interval == data_interval$interval, 
                                      data_interval$mean_step, 
                                      0))
#NA values in steps are replaced by mean steps at relevant interval
```

## New dataset
"data_na_filled" is the new dataset with NA values replaced

## New histogram, total steps, mean steps, median steps
```{r}
data_na_filled_sum <- data_na_filled %>% 
                      group_by(date) %>% 
                      summarise(sum_step = sum(steps))
hist(data_na_filled_sum$sum_step, xlab="Total number of steps", 
     main="Histogram of the total number of steps taken each day")
```

```{r, echo=FALSE}
print(paste0("mean total number of steps taken per day after imputing missing data ", round(mean(data_na_filled_sum$sum_step),2)))
print(paste0("median total number of steps taken per day after imputing missing data ", round(median(data_na_filled_sum$sum_step),2)))
# Filling missing data raises both total steps, mean steps, median steps compared to the previous dataset
```

# Are there differences in activity patterns between weekdays and weekends?
## Create a new factor variable in the dataset with two levels – “weekday” and “weekend”

```{r}
data_na_filled$date_type <- ifelse(tolower(weekdays(data_na_filled$date)) %in% c("saturday", "sunday"),
                                   "weekend", "weekday")
```

## Calculate mean and median of total steps by new levels of date type

```{r}
data_na_filled_sum_wkday <- data_na_filled %>% 
                            group_by(date_type, interval) %>% 
                            summarise(mean_step = mean(steps))
```

```{r}
library(lattice)
xyplot(mean_step ~ interval | date_type, data_na_filled_sum_wkday, 
       type="l", 
       lwd=1, 
       xlab="Interval", 
       ylab="Number of steps", 
       layout=c(1,2))
```

knit